---
title: "Additional figures in the paper for the three state formulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Additional figures in the paper for the three state formulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(MissingCases)
library(ggplot2)
```

## Convergence

We now show an example of the convergence of the system. We assume π= 0.848, α= 0.484,  γ= 0.772 and ϕ= 0.919. We chose a distribution for R to get some variability while keeping a relatively contained epidemic.

```{r}
set.seed(10)

p = 0.848
alpha = 0.484
gamma = 0.772
phi = 0.919

NGM <- define_NGM(p, alpha, gamma, phi)
EVNGM <- eigen(NGM)

Reff <- 1

n_t <- 80 #number of generations

Rt <- Reff/EVNGM$values[1] * runif(n=n_t, min = .2, max = 2)

X <- matrix(0,nrow = 3,ncol=n_t)
# Initialise number of cases in each group
X[,1] <- c(8,2,2)
# Simulate epidemic
for(i in 2:n_t){
  X[,i] <- Rt[i] * NGM %*% X[,i-1]
}
```

```{r, echo=FALSE, fig.cap="Figure 1: Number of cases for each generation of the example model where $p=0.46$, $\\alpha=0.68$, $\\gamma=0.47$, $\\phi = 0.60$ and $R_{N} \\sim \\mathcal{U}(0.2, 2)$"}
barplot(X[2,] + X[3,], ylab="Number of detected cases", xlab="Generations", col="#22329933")
```

```{r, echo=FALSE, fig.cap="Figure 2: Black dots are values derived from simulation while the red line shows the value derived analytically"}
plot(X[1,]/(X[1,]+X[2,]+X[3,]), ylim=c(0,1), ylab="Proportion of not detected cases", xlab="Number of generations", pch=20)
abline(h=EVNGM$vectors[1,1]/(EVNGM$vectors[1,1] + EVNGM$vectors[2,1] + EVNGM$vectors[3,1]), col='red')
```

We see that, as predicted by the analytic derivation, despite the unpredictable course of the epidemics (Figure 1), the proportion of infections that are not detected quickly reaches an equilibrium (Figure 2). This proportion at equilibrium can be calculated using the formula derived above. We note that the convergence is exponential and thus fast proving that λ_1≫λ_2 and λ_1≫λ_3. 


### Solution

We assume that π= 0.848, ϕ= 0.919 and get r_1= 0.25 and r_2= 0.20 from the NZ data. Using Equation 5, we estimate γ = 0.772. Solving the system gives the following solution: α= 0.484 and μ_ND1= 0.053 meaning that in this configuration, 5% of the infections are not detected and missing from the records. A graphical representation is given in Figure 3, where the solution of the system can be seen at the intersection of the two curves. No solution is found if the two curves do not intersect.

```{r}
p2 <- 0.848
gamma2 <- 0.772
r1 = 0.25
r2 =0.2
```

```{r, echo=FALSE, fig.width=6, fig.height=5, fig.cap="Figure 3: Derivation of $\\alpha$ the reduction in transmission due following tracing and $\\mu_{1}$ the proportion of not detected cases, assuming parameters above. The solid line links $\\mu$ and $\\alpha$ by using the proportion of de novo cases observed vs cases who are contact and followed up. The dashed line indicates the relationship given by the dynamics of the NGM."}
phi2 <- phi         #fix phi

#plot \mu as a function of alpha using the NGM
NGM3 <- matrix(0, nrow = 3, ncol = 3)

#Number of points to evaluate mu1 at
n_p <- 100
# Defines alpha vector
aa <- seq(from=0, to=1, length.out = n_p)
# Initialise vectors
mm_NGM3 <- rep(0, n_p)
mm_reap3 <- rep(0, n_p)

for(i in 1:n_p)
{
  #fix alpha
  alpha <- aa[i]
  #compute surveillance component of next generation matrix
  NGM3 <- define_NGM(p2, alpha, gamma2, phi2)
  # compute eigenvalues
  EVNGM3 <- eigen(NGM3)
  # compute proportion of missing cases using eigenvalue solution
  mm_NGM3[i] <- EVNGM3$vectors[1,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  # compute proportions of other cases
  mu_two <- EVNGM3$vectors[2,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  mu_three <- EVNGM3$vectors[3,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  # compute proportion of other cases using alternative relationship
  nu <- r2*phi2*gamma2/p2 - (1-phi2)
  mm_reap3[i] <- nu*(mu_two + alpha*mu_three)
}

dat = data.frame(x = c(aa,aa),
                 y = c(mm_reap3, mm_NGM3),
                 type = c(rep("Equation 7", length(mm_reap3)), rep("NGM", length(mm_NGM3))))

ggplot(dat) + geom_line(aes(x, y, linetype = type)) + 
  xlab("Reduction in transmission for cases who are being actively\n surveilled as contacts relative to other cases (" ~ alpha ~")") + ylab("Proportion of not detected cases (" ~ mu[ND] ~ ")") + 
   scale_x_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) + 
  theme_bw() + theme(legend.title = element_blank())
```


```{r, echo=FALSE}
#Use the find_sol function to find the values of alpha and \mu for the example
sol_example_3 <- compute_alpha_mu1(p=p2, phi=phi2, r1=r1, r2=r2)
print(sol_example_3)
```

```{r}
alpha2 <- 0.484
r1 = 0.25
r2 =0.2
p2 <- 0.848
gamma2 <- 0.772
```


```{r, echo=FALSE, fig.width=6, fig.height=5, fig.cap="Figure 4: Derivation of $\\phi$ the proportion of infected contacts recalled and $\\mu_{1}$ the proportion of not detected cases, assuming parameters above. The solid line links $\\mu$ and $\\phi$ by using the proportion of de novo cases observed vs cases who are contact and followed up. The dashed line indicates the relationship given by the dynamics of the NGM."}

#plot \mu as a function of alpha using the NGM
NGM3 <- matrix(0, nrow = 3, ncol = 3)

#Number of points to evaluate mu1 at
n_p <- 100
# Defines alpha vector
pp <- seq(from=0, to=1, length.out = n_p)
# Initialise vectors
mm_NGM3 <- rep(0, n_p)
mm_reap3 <- rep(0, n_p)

for(i in 1:n_p)
{
  #fix alpha
  phi <- pp[i]

  #compute surveillance component of next generation matrix
  NGM3 <- define_NGM(p2, alpha2, gamma2, phi)
  # compute eigenvalues
  EVNGM3 <- eigen(NGM3)
  # compute proportion of missing cases using eigenvalue solution
  mm_NGM3[i] <- EVNGM3$vectors[1,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  # compute proportions of other cases
  mu_two <- EVNGM3$vectors[2,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  mu_three <- EVNGM3$vectors[3,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  # compute proportion of other cases using alternative relationship
  nu <- r2*phi*gamma2/p2 - (1-phi)
  mm_reap3[i] <- nu*(mu_two + alpha2*mu_three)
}

dat = data.frame(x = c(pp,pp),
                 y = c(mm_reap3, mm_NGM3),
                 type = c(rep("Equation 7", length(mm_reap3)), rep("NGM", length(mm_NGM3))))

ggplot(dat) + geom_line(aes(x, y, linetype = type)) + 
  xlab("Proportion of infected contacts recalled (" ~ phi~ ")") + ylab("Proportion of not detected cases (" ~ mu[ND] ~ ")") + 
   scale_x_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) + 
  theme_bw() + theme(legend.title = element_blank())

```


```{r, echo=FALSE}
#Use the find_sol function to find the values of alpha and \mu for the example
sol_example_3 <- compute_phi_mu1(p=p2, alpha=alpha2, r1=r1, r2=r2)
print(sol_example_3)
```

```{r}
alpha2 <- 0.484
r1 = 0.25
r2 =0.2
phi2 = 0.919

```

```{r, echo=FALSE, fig.width=6, fig.height=5, fig.cap="Figure 5: Derivation of $\\pi$ the proportion of cases re-captured by community surveillance and $\\mu_{1}$ the proportion of not detected cases, assuming parameters above. The solid line links $\\mu$ and $\\pi$ by using the proportion of de novo cases observed vs cases who are contact and followed up. The dashed line indicates the relationship given by the dynamics of the NGM."}

#plot \mu as a function of alpha using the NGM
NGM3 <- matrix(0, nrow = 3, ncol = 3)

#Number of points to evaluate mu1 at
n_p <- 100
# Defines alpha vector
pp <- seq(from=0, to=1, length.out = n_p)
# Initialise vectors
mm_NGM3 <- rep(0, n_p)
mm_reap3 <- rep(0, n_p)

for(i in 1:n_p)
{
  #fix p
  p <- pp[i]
  gamma = p/(r1 + p)

  #compute surveillance component of next generation matrix
  NGM3 <- define_NGM(p, alpha2, gamma, phi2)
  # compute eigenvalues
  EVNGM3 <- eigen(NGM3)
  # compute proportion of missing cases using eigenvalue solution
  mm_NGM3[i] <- EVNGM3$vectors[1,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  # compute proportions of other cases
  mu_two <- EVNGM3$vectors[2,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  mu_three <- EVNGM3$vectors[3,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  # compute proportion of other cases using alternative relationship
  nu <- r2*phi2*gamma/p - (1-phi2)
  mm_reap3[i] <- nu*(mu_two + alpha2*mu_three)
}

dat = data.frame(x = c(pp,pp),
                 y = c(mm_reap3, mm_NGM3),
                 type = c(rep("Equation 7", length(mm_reap3)), rep("NGM", length(mm_NGM3))))

ggplot(dat) + geom_line(aes(x, y, linetype = type)) + 
  xlab("Proportion of cases cases detected in the community(" ~ pi~ ")") + ylab("Proportion of not detected cases (" ~ mu[ND] ~ ")") + 
   scale_x_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) + 
  theme_bw() + theme(legend.title = element_blank())

```


```{r, echo=FALSE}
#Use the find_sol function to find the values of alpha and \mu for the example
sol_example_3 <- compute_pi_mu1(phi=phi2, alpha=alpha2, r1=r1, r2=r2)
print(sol_example_3)
```

```{r, echo=FALSE, fig.width=6, fig.height=5, fig.cap="Figure 6: Derivation of $\\pi$ the proportion of cases re-captured by community surveillance and $\\mu_{1}$ the proportion of not detected cases, assuming parameters above. The solid line links $\\mu$ and $\\pi$ by using the proportion of de novo cases observed vs cases who are contact and followed up. The dashed line indicates the relationship given by the dynamics of the NGM."}

alpha2 <- 0.484
r1 = 0.25
r2 =0.2
phi2 = 0.7

#plot \mu as a function of alpha using the NGM
NGM3 <- matrix(0, nrow = 3, ncol = 3)

#Number of points to evaluate mu1 at
n_p <- 100
# Defines alpha vector
pp <- seq(from=0, to=1, length.out = n_p)
# Initialise vectors
mm_NGM3 <- rep(0, n_p)
mm_reap3 <- rep(0, n_p)

for(i in 1:n_p)
{
  #fix alpha
  p <- pp[i]
  gamma = p/(r1 + p)

  #compute surveillance component of next generation matrix
  NGM3 <- define_NGM(p, alpha2, gamma, phi2)
  # compute eigenvalues
  EVNGM3 <- eigen(NGM3)
  # compute proportion of missing cases using eigenvalue solution
  mm_NGM3[i] <- EVNGM3$vectors[1,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  # compute proportions of other cases
  mu_two <- EVNGM3$vectors[2,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  mu_three <- EVNGM3$vectors[3,1]/(EVNGM3$vectors[1,1] + EVNGM3$vectors[2,1] + EVNGM3$vectors[3,1])
  # compute proportion of other cases using alternative relationship
  nu <- r2*phi2*gamma/p - (1-phi2)
  mm_reap3[i] <- nu*(mu_two + alpha2*mu_three)
}

dat = data.frame(x = c(pp,pp),
                 y = c(mm_reap3, mm_NGM3),
                 type = c(rep("Equation 7", length(mm_reap3)), rep("NGM", length(mm_NGM3))))

ggplot(dat) + geom_line(aes(x, y, linetype = type)) + 
  xlab("Proportion of cases cases detected in the community(" ~ pi~ ")") + ylab("Proportion of not detected cases (" ~ mu[ND] ~ ")") + 
   scale_x_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) + 
  theme_bw() + theme(legend.title = element_blank())

```
